---
title: "time series analysis"
format: html
editor: visual
---

```{r}
#| echo: false

library(dplyr)
library(readr)
library(lme4)
library(lmerTest)
library(performance)
library(effectsize)
library(ggplot2)

source("99HF_functions.R")
source("99HF_random_forest.R")
set.seed(990)

```

# Data Cleaning

## Read in Data

```{r}

#models = c("claude", "gpt-4o", "deepseek-r1")
models = c("Llama-3.3-70B-Instruct", "Qwen2.5-72B-Instruct")

liwc_results_allmodels <- read_liwc_results_all(models)

liwc_results_allmodels |>
  group_by(model, scenario) |>
  summarize(
    n_aligned = sum(alignment == 1),
    n_not_aligned = sum(alignment == 0),
    pct_not_aligned = (n_not_aligned / n()),
    .groups = "drop"
  )

colnames(liwc_results_allmodels)
```

# Summary Statistics

```{r}

liwc_results_allmodels |>
  group_by(model, alignment) |>
  summarize(
    avg_tone = mean(Tone, na.rm = TRUE),
    .groups = "drop"
  )
```

# Lexical Variables Only

## Generate Variable Importance

```{r}

state_variable_importance <- list()

for (m in models) {
  
  print(m)
  liwc_results <- filter(liwc_results_allmodels, model == m)
  rf_input <- create_balanced_df_scenario(liwc_results)
  rf_input <- create_rf_data(rf_input, "state")
  state_variable_importance[[m]] = run_rf_model(rf_input)
  slice_head(state_variable_importance[[m]], n = 20) #show top 20 variables
  
}

View(state_variable_importance$claude)
View(state_variable_importance$`gpt-4o`)
View(state_variable_importance$`deepseek-r1`) 
```

## Logistic Regression

```{r}

state_standard_params <- list()
state_model <- list()

for (m in models) {

  print(m)
  
  #predictors <- setdiff(state_variable_importance[[m]]$LIWC_Variable[1:20], "")
  #predictors <- c("assent", "we", "Tone", "negate", "adj", "risk", "moral", "Clout", "assent_diff")
  #predictors <- c("assent", "we", "Tone", "negate", "adj", "risk", "moral", "i * you")
  predictors <- c("assent", "we", "negate", "adj", "risk", "assent_diff")
  
  formula_string <- paste("alignment ~", paste(predictors, collapse = " + "))

  state_model[[m]] <- glm(
    formula = as.formula(formula_string),
    data = filter(liwc_results_allmodels, model == m),
    family = binomial
  )

  print("Summary of Model")
  print(summary(state_model[[m]]))
  
  print("Predictive Power: AUC")
  print(performance_roc(state_model[[m]])) #AUC
  
  print("Check for Correlations: VIP")
  print(check_collinearity(state_model[[m]])) #VIP
  
  print("Effect Size: Standardized Parameters")
  state_standard_params[[m]] <- 
    standardize_parameters(state_model[[m]]) |> #effect size 
    as.data.frame() |>
    mutate(abs_coef = abs(Std_Coefficient)) |>
    arrange(desc(abs_coef))

}

View(state_standard_params$claude)
View(state_standard_params$`gpt-4o`)
View(state_standard_params$`deepseek-r1`) 
```

## Combining models together

```{r}

m = "all"

rf_input <- create_balanced_df_scenario(liwc_results_allmodels)
rf_input <- create_rf_data(rf_input, "state")
state_variable_importance[[m]] = run_rf_model(rf_input)
slice_head(state_variable_importance[[m]], n = 20) #show top 20 variables

```

```{r}

# Logistic Regression

#predictors <- setdiff(state_variable_importance[[m]]$LIWC_Variable[1:20], "")
predictors <- setdiff(state_standard_params$all$Parameter[2:11], "")
#predictors <- c("assent", "we", "Tone", "negate", "adj", "risk", "moral", "Clout")
  
formula_string <- paste("alignment ~", paste(predictors, collapse = " + "))

state_model[[m]] <- glm(
    formula = as.formula(formula_string),
    data = liwc_results_allmodels,
    family = binomial
  )

  print("Summary of Model")
  print(summary(state_model[[m]]))
  
  print("Predictive Power: AUC")
  print(performance_roc(state_model[[m]])) #AUC
  
  print("Check for Correlations: VIP")
  print(check_collinearity(state_model[[m]])) #VIP
  
  print("Effect Size: Standardized Parameters")
  state_standard_params[[m]] <- 
    standardize_parameters(state_model[[m]]) |> #effect size 
    as.data.frame() |>
    mutate(abs_coef = abs(Std_Coefficient)) |>
    arrange(desc(abs_coef))
  
View(state_standard_params$`all`) 
```

# Delta Variables Only

## Generate Variable Importance

```{r}
delta_variable_importance <- list()

for (m in models) {
  
  print(m)
  liwc_results <- filter(liwc_results_allmodels, model == m)
  rf_input <- create_balanced_df_scenario(liwc_results)
  rf_input <- create_rf_data(rf_input, "delta")
  delta_variable_importance[[m]] = run_rf_model(rf_input)
  slice_head(delta_variable_importance[[m]], n = 20) #show top 20 variables
  
}

View(delta_variable_importance$claude)
View(delta_variable_importance$`gpt-4o`)
View(delta_variable_importance$`deepseek-r1`) 
```

## Logistic Regression

```{r}

delta_standard_params <- list()
delta_model <- list()

for (m in models) {

  print(m)
  
  predictors <- setdiff(delta_variable_importance[[m]]$LIWC_Variable[1:20], "")
  #predictors <- c("assent", "we", "Tone", "negate", "adj", "risk", "moral", "Clout")
  
  formula_string <- paste("alignment ~", paste(predictors, collapse = " + "))

  delta_model[[m]] <- glm(
    formula = as.formula(formula_string),
    data = filter(liwc_results_allmodels, model == m),
    family = binomial
  )

  print("Summary of Model")
  print(summary(delta_model[[m]]))
  
  print("Predictive Power: AUC")
  print(performance_roc(delta_model[[m]])) #AUC
  
  print("Check for Correlations: VIP")
  print(check_collinearity(delta_model[[m]])) #VIP
  
  print("Effect Size: Standardized Parameters")
  delta_standard_params[[m]] <- 
    standardize_parameters(delta_model[[m]]) |> #effect size 
    as.data.frame() |>
    mutate(abs_coef = abs(Std_Coefficient)) |>
    arrange(desc(abs_coef))

}

View(delta_standard_params$claude)
View(delta_standard_params$`gpt-4o`)
View(delta_standard_params$`deepseek-r1`) 
```

# Combined - Absolute & Delta Data Together

## Generate Variable Importance

```{r}

combined_variable_importance <- list()

for (m in models) {
  
  print(m)
  liwc_results <- filter(liwc_results_allmodels, model == m)
  rf_input <- create_balanced_df_scenario(liwc_results)
  rf_input <- create_rf_data(rf_input, "combined")
  combined_variable_importance[[m]] = run_rf_model(rf_input)
  slice_head(combined_variable_importance[[m]], n = 20) #show top 20 variables
  
}

View(combined_variable_importance$claude)
View(combined_variable_importance$`gpt-4o`)
View(combined_variable_importance$`deepseek-r1`)
```

## Logistic Regression

```{r}
combined_standard_params <- list()
combined_model <- list()

for (m in models) {

  print(m)
  
  #predictors <- setdiff(combined_variable_importance[[m]]$LIWC_Variable[1:20], "")
  predictors <- c("assent", "we", "Tone", "negate", "adj", "risk", "moral", "Clout", "assent_diff")
  #predictors <- c("assent", "assent_diff", "we", "negate", "adj", "risk", "we_diff")
  
  formula_string <- paste("alignment ~", paste(predictors, collapse = " + "))

  combined_model[[m]] <- glm(
    formula = as.formula(formula_string),
    data = filter(liwc_results_allmodels, model == m),
    family = binomial
  )

  print("Summary of Model")
  print(summary(combined_model[[m]]))
  
  print("Predictive Power: AUC")
  print(performance_roc(combined_model[[m]])) #AUC
  
  print("Check for Correlations: VIP")
  print(check_collinearity(combined_model[[m]])) #VIP
  
  print("Effect Size: Standardized Parameters")
  combined_standard_params[[m]] <- 
    standardize_parameters(combined_model[[m]]) |> #effect size 
    as.data.frame() |>
    mutate(abs_coef = abs(Std_Coefficient)) |>
    arrange(desc(abs_coef))

}

View(combined_standard_params$claude)
View(combined_standard_params$`gpt-4o`)
View(combined_standard_params$`deepseek-r1`) 
```

## Combining models together

```{r}

# Logistic Regression

#m = "all"
#m = "claude"
#m = "gpt-4o"
#m = "deepseek-r1"
m = "Llama-3.3-70B-Instruct"
#m = "Qwen2.5-72B-Instruct"

#predictors <- setdiff(state_variable_importance[[m]]$LIWC_Variable[1:20], "")
#predictors <- setdiff(state_standard_params$all$Parameter[2:11], "")
predictors <- c("assent", "we", "Tone", "negate", "adj", "risk", "moral", "Clout", "assent_diff")
  
formula_string <- paste("alignment ~", paste(predictors, collapse = " + "))

state_model[[m]] <- glm(
    formula = as.formula(formula_string),
    data = liwc_results_allmodels,
    family = binomial
  )

  print("Summary of Model")
  print(summary(state_model[[m]]))
  
  print("Predictive Power: AUC")
  print(performance_roc(state_model[[m]])) #AUC
  
  print("Check for Correlations: VIP")
  print(check_collinearity(state_model[[m]])) #VIP
  
  print("Effect Size: Standardized Parameters")
  state_standard_params[[m]] <- 
    standardize_parameters(state_model[[m]]) |> #effect size 
    as.data.frame() |>
    mutate(abs_coef = abs(Std_Coefficient)) |>
    arrange(desc(abs_coef))
  
View(state_standard_params$`all`) 
```

# ASSENT OVER TIME

## Create new turn_before_switch variable

```{r}

#MODEL
#m = "claude"
#m = "gpt-4o"
m = "deepseek-r1"
#m = "Llama-3.3-70B-Instruct"
#m = "Qwen2.5-72B-Instruct"
liwc_results <- filter(liwc_results_allmodels, model == m)

# Standardizing the timeline: Aligning conversations by counting backward 
# from the exact turn where the opinion switch occurred (Turn 0)
plot_data <- liwc_results |>
  group_by(question_id, prompt) |>
  mutate(
    turns_before_switch = turn_no - ToF - 1,
    #ToF = ifelse(ToF == 5, "NA", ToF),
    turn_of_flip = factor(
      ifelse(ToF == 5, "NA", ToF + 1),
      levels = c("NA", "2", "3", "4", "5"),
      labels = c("No Flip", "Turn 2", "Turn 3", "Turn 4", "Turn 5")
      ),
  ) |>
  filter(turns_before_switch >= -5 & turns_before_switch <= 0) |> # Focus on the critical window
  filter(ToF > 0)

plot_data |>
  group_by(turn_of_flip) |>
  summarize(
    n = n()
  ) 
```

## Visualize in a Plot

```{r}

# Plotting the trajectory
ggplot(plot_data, aes(x = turns_before_switch, y = assent, group = turn_of_flip, color = turn_of_flip)) +
  stat_summary(fun = mean, geom = "line", size = 1.2) + # Average trajectory path
  #stat_smooth(method = "lm", formula = y ~ x, se = FALSE, linetype = "dashed", size = 1.3) +
  stat_summary(fun.data = mean_se, geom = "ribbon", alpha = 0.15, aes(fill = turn_of_flip)) + # Error bands
  theme_minimal() +
  labs(
    title = paste0("Trajectory of Assent Prior to Stance Switch ","[", m, "]"),
    x = "Turns Relative to Switch Point (Turn 0)",
    y = "Mean LIWC Assent Score"
  )
```

```{r}

plot_data$turn_of_flip <- relevel(plot_data$turn_of_flip, ref = "No Flip")

multi_slope_model <- glm(assent ~ turn_no * turn_of_flip, data = plot_data)
summary(multi_slope_model)
```

Summary Statistics
